Perturbative PDF of the total magnetization of the 4D Ising model

### 2020-04-19 paper/paper.tex
Proofread

\title{Perturbative probability distribution of the average Ising magnetization
in 4 dimensions.}

@@

\abstract{We compute, at one loop in perturbation theory, the probability density
function of the average magnetization $\bar{\phi}$ of the Ising model, on both
the 4-torus and the 4-sphere. The perturbative expansion covers the entire
symmetric phase and a neighborhood of the critical point. We find that, at the
critical point, the PDF has the asymptotic behavior $p(\bar{\phi})\sim
\exp(-c \bar{\phi}^4)$, and hence the critical value of the Binder cumulant
is $U = 1 - \frac{4\,\Gamma(5/4)^2}{ 3\,\Gamma(3/4)^2}$. We validate our
results by comparison with Monte Carlo simulation.}
\begin{document}

@@

\section{Main results}

The average magnetization $\bar{\phi} = \frac{1}{V} \sum_{i} \sigma_i$ of the
Ising model is the simplest observable in Monte Carlo simulations. When the
system size $L$ is large, the probability distribution of $\bar{\phi}$ changes
qualitatively between the two phases of the model. In the symmetric phase,
\textit{i. e.} when the correlation length $\xi$ is large in lattice units, but
small compared to the system size, the distribution is well approximated by a
zero-mean normal. In the broken symmetry phase, still for $1 \ll \xi \ll L$,
the distribution is well approximated by a mixture of two normals centered at
non-zero values $\pm \bar{\phi}_0$.

Near the critical point, \textit{ie.} $1 \ll L \ll \xi$, the probability
distribution of $\bar{\phi}$ is a non-trivial function of $\bar{\phi}$, at
least in two and three dimensions.  It is not immediately clear if the same is
also true in four dimensions. Since the field theory that describes the Ising
critical point in four dimensions becomes weakly coupled at low energy, one may
expect the probability distribution of $\bar{\phi}$ to remain a zero-centered
normal at the critical point as well as in the symmetric phase.

To clarify this issue, we compute the logarithm of the PDF of $\bar{\phi}$ at
one-loop in $4$ dimensions, and show that, for very large system size, the
critical probability distribution is not normal, but rather of the form $-\log
p(\bar{\phi})\sim f(L) \bar{\phi}^4$. We find that corrections to this form
vanish very slowly with increasing system size, like $1/\log L$.

@@

Thus we consider the $\phi^4$ field theory with action:
\begin{equation}
  S[\phi] \equiv \int \dd^4 x\ \left(
    \frac{1}{2} \phi(x)\left(\Delta + r_0\right)\phi(x) +
    \frac{1}{24} u_0 \phi(x)^4\right)
\end{equation}
defined on a 4-torus of radius $R$, i.e. $x_i\sim x_i + 2\pi R$. In this field
theory context, the PDF of the average magnetization is defined as:
\begin{equation}
    \label{eq:pdf_definition}
    p(\bar\phi) \equiv \frac{1}{Z} \int \left[\mathcal{D}\phi\right]
    \delta\left(\bar{\phi} - \frac{1}{V}\int\dd^4 x\ \phi(x)\right)
    e^{-S[\phi]}\,,
\end{equation}
and at one loop it evaluates to:
\begin{equation}
\label{eq:log_pdf}
\begin{split}
    -\log p(\bar{\phi}) = \mathrm{const.} +  V\Bigg(
  &\frac{1}{2} \frac{\bar{\phi}^2}{R^2} \left(
    r R^2 + \frac{g}{6} \left(f_1\left(r R^2\right) - 1\right)
    + O\left(g^2\right)\right) + \\
  & \frac{2\pi^2}{9}g\bar{\phi}^4\left(
    1 - \frac{g}{2}f_2\left(r R^2\right)
    + O\left(g^2\right)\right) + \\
  & \frac{8\pi^4}{81} g^3 \bar{\phi}^6 R^2\left(
    f_3\left(r R^2\right)
    + O\left(g\right)\right) + O\left(g^4\bar{\phi}^8\right) \Bigg)\,,
\end{split}
\end{equation}
where the coefficients $r$ and $g$ are the renormalized counterparts to $r_0$
and $u_0$. An explicit expression for the functions $f_1$, $f_2$ and $f_3$ is
reported in section \ref{sect:derivation} (\ref{eq:f_functions}), and they are
plotted in
fig.~\ref{fig:f_functions}.

@@

Eq. (\ref{eq:log_pdf}) is our central result, and we now describe it
at length. We first illustrate its domain of validity and its dependence on the
renormalized couplings $r$, $g$ on a torus of fixed radius $R$, then describe
its dependence on $R$ at fixed bare couplings $r_0$, $u_0$.

The dimension-2 coupling $r$ controls the cross-over between the symmetric and
broken-symmetry phases. The expression (\ref{eq:log_pdf}) is valid for all
positive values of $r$, and also for negative values, provided that $rR^2
\gtrsim -1/2$.  For lower values of $r$, additional modes of the Laplacian
condense beside the zero-mode, and this invalidates the perturbative expansion.
This becomes manifest in a divergence of the functions $f_i$ as their argument
approaches -1.  On the other hand, $f_1$, $f_2$, $f_3$ all go to zero as their
argument approaches positive infinity, and hence for large $r$ the probability
distribution of $\bar{\phi}$ is a zero-centered normal:
\begin{equation}
    p(\bar{\phi}) \sim N e^{- \frac{1}{2} V r \bar{\phi}^2}
  \quad \text{for} \quad
  rR^2 \gg 1\,.
\end{equation}

The dimensionless coupling $g$ is the parameter of the perturbative expansion.
The expansion is valid for $g \ll 1$, provided $rR^2$ is sufficiently far from
the bound discussed above.

In order to describe how $p(\bar{\phi})$ depends on $R$ at fixed bare couplings
$r_0$, $u_0$, it is necessary to account for renormalization effects. In
general, the renormalized couplings $r$, $g$ depend on a renormalization scale
$\mu$, and it is possible to derive an expression for $p(\bar{\phi})$ evaluated
at a generic value of $\mu$. However, for simplicity,
(\ref{eq:log_pdf}) is evaluated at the scale $\mu^2 = r + R^{-2}$.
This choice of $\mu$ is optimal for the reliability of perturbation theory,
because it avoids the emergence of large logarithms over the widest possible
range of parameters.

The renormalization group equations for $r$ and $g$ are:
\begin{equation}
  \label{eq:renormalization_group}
  \mu \frac{\dd g}{\dd \mu} = g^2\,,\quad
  \mu \frac{\dd r}{\dd \mu} = \frac{1}{3}g r\,.
\end{equation}

These can be integrated and combined with the condition $\mu^2 = r + R^{-2}$ to
obtain a system of equations\footnote{There are of course many alternative,
arguably simpler, solutions that differ by sub-leading orders in an expansion in
$g(R_1)$. The one displayed here is the exact solution to
(\ref{eq:renormalization_group}).} connecting the renormalized couplings at two
different values of $R$:
\begin{equation}
  \frac{g(R_1)}{g(R_2)} = \frac{r(R_1)^3}{r(R_2)^3} = 
  1 - \frac{1}{2} g(R_1)\log\frac{r(R_2) + R_2^{-2}}{r(R_1) + R_1^{-2}}\,.
\end{equation}

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{renormalization_group.png}
\end{center}
    \caption{\label{fig:renormalization_group} A few solutions to
    (\ref{eq:renormalization_group}) with $g(R_1) = 0.45$.  In the left plot,
    the shaded region shows where the perturbative expansion becomes invalid.
    Similarly in the right plot, the lines become dotted outside of the
    perturbative region. Note how the sign of $r$ is preserved, $r = 0$ is a
    solution, and $g$ vanishes for $R_2 \gg R_1$.}
\end{figure}

A few solutions to these equations are shown in
fig.~\ref{fig:renormalization_group}. Within the perturbative regime $g \ll 1$,
$r R^2 \gtrsim -1/2$, the coupling $r$ varies very little with $R$, and, as is
clear from the differential form (\ref{eq:renormalization_group}), the sign of
$r$ is always preserved. Thus we conclude that the critical point is at $r
= 0$, and the symmetric phase is realized for $r > 0$.

At the critical point, the renormalized coupling $g$ follows the simplified renormalization group equation:
\begin{equation}
    \frac{1}{g(R_2)} - \frac{1}{g(R_1)} = \log\frac{R_2}{R_1}\,,
\end{equation}
and hence, as system size grows, the renormalized coupling $g$ goes to zero.
In this regime, the quartic term in (\ref{eq:log_pdf}) dominates all the
others. This is perhaps most evident if the PDF is expressed in terms of the
rescaled quantity $\bar{\varphi} = g^{\frac{1}{4}} \bar{\phi}$, whose
variance remains finite as $g\to0$.

Thus we conclude that, at the critical point, for sufficiently large system
size:
\begin{equation}
    \label{eq:log_pdf_critical_point}
    -\log p(\bar{\phi}) \sim \mathrm{const} + \frac{2\pi^2}{9} V g\bar{\phi}^4\,,
\end{equation}
with sub-leading terms vanishing like $1 / \log(R)$.

In Monte Carlo simulations, the qualitative behavior of the distribution of the
magnetization is often characterized by measuring the so-called Binder
cumulant \cite{Binder1981}:
\begin{equation}
    U = 1 - \frac{\left\langle \bar \phi^4 \right \rangle}
    {3\left\langle \bar \phi^2 \right \rangle^2},
\end{equation}
which is constructed to be independent of the overall scale of $\bar{\phi}$,
and to be zero if $\bar{\phi}$ is normally distributed. From
(\ref{eq:log_pdf_critical_point}) we conclude that, on a 4-Torus, at the
critical point\footnote{ Incidentally, we note that this result is not in
contradiction with the claim in the original paper \cite{Binder1981} that the
critical Binder cumulant in 4 dimensions is zero, because in the original paper
the cumulant is not computed from the total magnetization, but rather from the
average magnetization of blocks of spins much smaller than the overall system
size.}:
\begin{equation}
    \label{eq:critical_binder_cumulant}
    U = 1 - \frac{4\,\Gamma\left(\frac{5}{4}\right)^2}
    {3\,\Gamma\left(\frac{3}{4}\right)^2} 
    = 0.27052\ldots
\end{equation}

In fig.~\ref{fig:binder_cumulant}, we show a comparison of the Binder cumulant
computed from (\ref{eq:log_pdf}) near the critical point, and the results of
Monte Carlo simulation of the $4D$ Ising model. The agreement is excellent except
for the smallest system size $L = 8$, probably because of contributions from
irrelevant operators. Notice how slowly the finite size Binder cumulant
approaches the asymptotic value (\ref{eq:critical_binder_cumulant}).

@@

We now describe briefly how the result (\ref{eq:log_pdf}) is obtained. The
perturbative approach is very similar to the computation of the effective
action, as in \textit{e.g.} \cite{WeinbergEffectiveAction}, except that we are
interested in the whole probability distribution of $\bar{\phi}$, instead of
just the expected value. The main difficulty lies in evaluating the loop
integrals at finite size.

First of all we observe that the average magnetization
$\bar{\phi}$ is proportional to the zero momentum mode of the field $\phi$:
\begin{equation}
    \phi(x) = \frac{1}{V} \sum_{n \in \mathbb{Z}^4} \phi_{n} 
    e^{i \frac{n\cdot x}{R}}\,;\quad \bar{\phi} = \frac{1}{V}\int\dd^4 x\ \phi(x) = \frac{1}{V}\phi_{n = 0}\,.
\end{equation}

Thus, because of the delta function in (\ref{eq:pdf_definition}), the component
$\phi_{0}$ becomes an external field, whereas the other components $\phi_{n}$
are still part of the functional integral. Separating the zero momentum mode
from the other modes in the action yields:
\begin{equation}
\begin{split}
    S(\bar{\phi}, \phi) =& 
    V \left(\frac{1}{2} r_0 \bar{\phi}^2 + \frac{u_0}{24} \bar{\phi}^4\right) +
    \frac{1}{2V} \sum_{n} \left(\frac{n^2}{R^2} + r_0 +
      \frac{1}{2} u_0\bar{\phi}^2 \right)\phi_{n} \phi_{-n} +\\
    &\frac{u_0}{6V^2} \bar{\phi} 
        \sum_{n_1, n_2} \phi_{n_1} \phi_{n_2} \phi_{-n_1 -n_2} +
    \frac{u_0}{24V^3} \sum_{n_1, n_2, n_3} \phi_{n_1} \phi_{n_2}
        \phi_{n_3} \phi_{-n_1 -n_2 -n_3}\,,
\end{split}
\end{equation}
where all summations now are over $\mathbb{Z}^4 \setminus \{0\}$.

We also introduce renormalized couplings
\begin{equation}
    u_0 = u(1 + u\delta u)\,;\quad r_0 = r(1 + u \delta r)\,,
\end{equation}
and we obtain the following edges and vertices in the diagrammatic expansion:
\begin{displaymath}
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[thick] (-1, 0) -- (1, 0);
    \node[left] at (-1, 0) {$n_1$};
    \node[right] at (1, 0) {$n_2$};
\end{tikzpicture}
    \quad
    \frac{V}{n_1^2R^{-2} + r}\delta_{n_1 + n_2}
\quad\quad
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[densely dashed, thick] (-1, -1) -- (0, 0) -- (-1, 1);
    \draw[thick] (+1, -1) -- (0, 0) -- (+1, 1);
    \node[above right] at (1, 1) {$n_1$};
    \node[below right] at (1, -1) {$n_2$};
\end{tikzpicture}
    -\frac{u\bar{\phi}^2}{4V}\delta_{n_1 + n_2}
\end{displaymath}
\begin{displaymath}
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[densely dashed, thick] (-1, -1) -- (0, 0);
    \draw[thick] (0, 0) -- (-1, 1);
    \draw[thick] (+1, -1) -- (0, 0) -- (+1, 1);
    \node[above left] at (-1, 1) {$n_1$};
    \node[above right] at (1, 1) {$n_2$};
    \node[below right] at (1, -1) {$n_3$};
\end{tikzpicture}
    -\frac{u\bar{\phi}}{6V^2}\delta_{n_1 + n_2 + n_3}
\quad\quad
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[thick] (-1, -1) -- (1, 1);
    \draw[thick] (-1, 1) -- ( 1, -1);
    \node[above left] at (-1, 1) {$n_1$};
    \node[above right] at (1, 1) {$n_2$};
    \node[below right] at (1, -1) {$n_3$};
    \node[below left] at (-1, -1) {$n_4$};
\end{tikzpicture}
    -\frac{u}{24V^3}\delta_{n_1 + n_2 + n_3 + n_4}
\end{displaymath}
plus additional vertices associated with the counterterms $\delta r$, $\delta
u$ which we do not list for brevity.

>>>>

\title{Perturbative PDF of the total magnetization of the 4D Ising model }

@@

\abstract{We compute, at one loop in perturbation theory, the probability
density function of the total magnetization $M$ of the Ising model, on both the
4-torus and the 4-sphere. We develop a single perturbative expansion that is
valid both in the symmetric phase, as long as the correlation length $\xi$ is
large in lattice units, and in the broken symmetry phase, as long as the
correlation length is large compared to the system size $L$. We find that, at
the critical point, for very large system size in lattice units, the PDF
approaches $p(M)\sim \exp(-f(L) M^4)$, and hence the critical value of the
Binder cumulant of the total magnetization is $U = 1 -
\frac{4\,\Gamma(5/4)^2}{3\,\Gamma(3/4)^2}$. We validate our results by
comparison with Monte Carlo simulation.}
\begin{document}

@@

\section{Main result}

When the system size $L$ is large, the probability distribution of the total
magnetization $M$ of the Ising model changes qualitatively between the two
phases of the model. In the symmetric phase, \textit{i.e.} when the correlation
length $\xi$ is large in lattice units, but small compared to the system size,
the distribution is well approximated by a zero-mean normal.  In the broken
symmetry phase, still for $1 \ll \xi \ll L$, the distribution is well
approximated by a mixture of two normals centered at non-zero values $\pm
M_0$.

Near the critical point $1 \ll L \ll \xi$, in two and three dimensions, the
probability distribution of $M$ is a non-trivial function of $M$.  It is not
immediately clear if the same is also true in four dimensions. Since the field
theory that describes the Ising critical point in four dimensions becomes
weakly coupled at low energy, one may expect the probability distribution of
$M$ to remain a zero-centered normal even at the critical point. This is in
fact the case for the total magnetization of a subsystem of intermediate size:
large in lattice units, but small compared to $L$ \cite{Binder1981}. Here we
show that the same is not true for the total magnetization of the entire
system. We compute the logarithm of the PDF of $M$ at one-loop in $4$
dimensions, and show that, for very large system size, the critical probability
distribution is not normal, but rather of the form $-\log p(M)\sim f(L) M^4$.
We find that corrections to this form vanish very slowly with increasing system
size, like $1/\log L$.

@@

When the correlation length is large in lattice unit, the Ising model in 4
dimensions is described by the $\phi^4$ field theory:
\begin{equation}
    S[\phi] \equiv \int_{\mathcal{M}} \left(
    \frac{1}{2} \phi\left(\Delta + r_0\right)\phi +
    \frac{1}{24} u_0 \phi^4\right)\,.
\end{equation}

Here we take the manifold $\mathcal{M}$ to be a 4-torus of radius $R$,
\textit{i.e.} $x_i\sim x_i + 2\pi R$. In this field theory context, we define
the PDF of the average magnetization $m = M / V$ as:
\begin{equation}
    \label{eq:pdf_definition}
    p(m) \equiv \frac{1}{Z} \int \left[\mathcal{D}\phi\right]
    \delta\left(m - \frac{1}{V}\int\dd^4 x\ \phi(x)\right)
    e^{-S[\phi]}\,,
\end{equation}
and we evaluate it at one loop in perturbation theory, obtaining:
\begin{equation}
\label{eq:log_pdf}
\begin{split}
    -\log p(m) = \mathrm{const.} +  V\Bigg(
  &\frac{1}{2} \frac{m^2}{R^2} \left(
    r R^2 + \frac{g}{6} \left(f_1\left(r R^2\right) - 1\right)
    + O\left(g^2\right)\right) + \\
  & \frac{2\pi^2}{9}gm^4\left(
    1 - \frac{g}{2}f_2\left(r R^2\right)
    + O\left(g^2\right)\right) + \\
  & \frac{8\pi^4}{81} g^3 m^6 R^2\left(
    f_3\left(r R^2\right)
    + O\left(g\right)\right) + O\left(g^4m^8\right) \Bigg)\,.
\end{split}
\end{equation}
Here the coefficients $r$ and $g$ are the renormalized counterparts to $r_0$
and $u_0$; the functions $f_1$, $f_2$ and $f_3$ are plotted in
fig.~\ref{fig:f_functions}, and an explicit expression (\ref{eq:f_functions})
is given below.

@@

This is our central result, and we now discuss it at length. We first
illustrate its domain of validity and its dependence on the renormalized
couplings $r$, $g$ on a torus of fixed radius $R$, then describe its dependence
on $R$ at fixed bare couplings $r_0$, $u_0$.

The dimension-2 coupling $r$ controls the cross-over between the symmetric and
broken-symmetry phases. The expression (\ref{eq:log_pdf}) is valid for  $rR^2
\gtrsim -1/2$.  For lower values of $r$, the perturbative vacuum becomes
unstable, invalidating the perturbative expansion. For this reason, the
functions $f_i$ diverge as their argument approaches -1.  On the other hand,
$f_1$, $f_2$, $f_3$ all go to zero as their argument approaches positive
infinity, and hence for large $r$ the probability distribution of $m$ is a
zero-centered normal:
\begin{equation}
    p(m) \sim N e^{- \frac{1}{2} V r m^2}
  \quad \text{for} \quad
  rR^2 \gg 1\,.
\end{equation}

The dimensionless coupling $g$ is the parameter of the perturbative expansion.
The expansion is valid for $g \lesssim 1$, provided $rR^2$ is sufficiently far
from the bound discussed above.

In order to describe how $p(m)$ depends on $R$ at fixed bare couplings $r_0$,
$u_0$, it is necessary to account for renormalization effects. The specific
renormalization scheme we used is described in section \ref{sect:derivation},
and (\ref{eq:log_pdf_with_mu}) gives expression for $p(m)$ evaluated at a
generic renormalization scale $\mu$. However, for simplicity, we chose to
evaluate (\ref{eq:log_pdf}) at the scale $\mu^2 = r + R^{-2}$.  This choice of
$\mu$ is optimal for the reliability of perturbation theory, because it avoids
the emergence of large logarithms over the widest possible range of parameters.

At one loop, the Callan-Symanzik equations for $r$ and $g$ are:
\begin{equation}
  \label{eq:callan_symanzik}
  \mu \frac{\dd g}{\dd \mu} = g^2\,,\quad
  \mu \frac{\dd r}{\dd \mu} = \frac{1}{3}g r\,.
\end{equation}

These can be integrated and combined with the condition $\mu^2 = r + R^{-2}$ to
obtain a system of equations\footnote{There are of course many alternative,
arguably simpler, solutions that differ by sub-leading orders in an expansion in
$g(R_1)$. The one displayed here is the exact solution to
(\ref{eq:callan_symanzik}).} connecting the renormalized couplings at two
different values of $R$:
\begin{equation}
  \label{eq:renormalized_relation}
  \frac{g(R_1)}{g(R_2)} = \frac{r(R_1)^3}{r(R_2)^3} = 
  1 - \frac{1}{2} g(R_1)\log\frac{r(R_2) + R_2^{-2}}{r(R_1) + R_1^{-2}}\,.
\end{equation}

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{renormalization_group.png}
\end{center}
    \caption{\label{fig:renormalization_group} A few solutions to
    (\ref{eq:renormalized_relation}) with $g(R_1) = 0.45$.  In the left plot,
    the shaded region shows where the perturbative vacuum becomes unstable,
    invalidating the perturbative expansion. Similarly, in the right plot, the
    lines become dotted outside of the perturbative region. Note how the sign
    of $r$ is preserved and $r = 0$ is a solution.}
\end{figure}

A few solutions to this system of equations are shown in
fig.~\ref{fig:renormalization_group}. Within the perturbative regime $g
\lesssim 1$, $r R^2 \gtrsim -1/2$, the coupling $r$ varies little with
$R$, and, as is clear from the differential form (\ref{eq:callan_symanzik}),
the sign of $r$ is always preserved. Thus we conclude that the critical point
is at $r = 0$, and the symmetric phase is realized for $r > 0$.

At the critical point, the renormalized coupling $g$ follows the simplified
Callan-Symanzik equation
\begin{equation}
    \frac{1}{g(R_2)} - \frac{1}{g(R_1)} = \log\frac{R_2}{R_1}\,,
\end{equation}
and hence, as system size grows, the renormalized coupling $g$ goes to zero as
$1 / \log(R)$.  In this regime, the quartic term in (\ref{eq:log_pdf})
dominates all the others. This is perhaps most evident if the PDF is expressed
in terms of the rescaled quantity $\bar{m} = g^{\frac{1}{4}} m$, whose variance
remains finite as $g\to0$. Thus we conclude that, at the critical point, for
sufficiently large system size:
\begin{equation}
    \label{eq:log_pdf_critical_point}
    -\log p(m) \sim \mathrm{const} + \frac{2\pi^2}{9} V gm^4\,.
\end{equation}

In Monte Carlo simulations, the qualitative behavior of the distribution of the
magnetization is often characterized by measuring the so-called Binder
cumulant \cite{Binder1981}:
\begin{equation}
    \label{eq:binder_cumulant}
    U = 1 - \frac{\left\langle m^4 \right \rangle}
    {3\left\langle m^2 \right \rangle^2}\,.
\end{equation}

This quantity is constructed to be independent of the overall scale of $m$,
and to be zero if $m$ is normally distributed. From
(\ref{eq:log_pdf_critical_point}) we conclude that, on a 4-Torus, at the
critical point:
\begin{equation}
    \label{eq:critical_binder_cumulant}
    U = 1 - \frac{4\,\Gamma\left(\frac{5}{4}\right)^2}
    {3\,\Gamma\left(\frac{3}{4}\right)^2} 
    = 0.27052\ldots
\end{equation}

In fig.~\ref{fig:binder_cumulant}, we show a comparison of the Binder cumulant
computed from (\ref{eq:log_pdf}) near the critical point, and the results of
Monte Carlo simulation of the $4D$ Ising model. The agreement is excellent
except for the smallest system size $L = 8$. Notice how slowly the finite size
Binder cumulant approaches the asymptotic value
(\ref{eq:critical_binder_cumulant}).

@@

We now describe briefly how the result (\ref{eq:log_pdf}) is obtained. The
perturbative approach is similar to the computation of the effective
action, as in \textit{e.g.} \cite{WeinbergEffectiveAction}, except that we are
interested in the whole probability distribution of $m$, instead of
just the expected value. The main difficulty lies in evaluating the loop
integrals at finite size.

The average magnetization $m$ is proportional to the zero-momentum mode of the
field:
\begin{equation}
    \phi(x) = \frac{1}{V} \sum_{n \in \mathbb{Z}^4} \phi_{n} 
    e^{i \frac{n\cdot x}{R}}\,;\quad m = \frac{1}{V}\int\dd^4 x\ \phi(x) = \frac{1}{V}\phi_{n = 0}\,.
\end{equation}

Because of the delta function in (\ref{eq:pdf_definition}), the zero-mode
becomes an external field, whereas all other modes are still part of the
functional integral. Separating the zero-mode from the other modes in the
action yields:
\begin{equation}
\begin{split}
    S(m, \phi) =& 
    V \left(\frac{1}{2} r_0 m^2 + \frac{u_0}{24} m^4\right) +
    \frac{1}{2V} \sum_{n} \left(\frac{n^2}{R^2} + r_0 +
      \frac{1}{2} u_0m^2 \right)\phi_{n} \phi_{-n} +\\
    &\frac{u_0}{6V^2} m 
        \sum_{n_1, n_2} \phi_{n_1} \phi_{n_2} \phi_{-n_1 -n_2} +
    \frac{u_0}{24V^3} \sum_{n_1, n_2, n_3} \phi_{n_1} \phi_{n_2}
        \phi_{n_3} \phi_{-n_1 -n_2 -n_3}\,,
\end{split}
\end{equation}
where all summations now are over $\mathbb{Z}^4 \setminus \{0\}$.

We introduce renormalized couplings
\begin{equation}
    u_0 = u(1 + u\delta u)\,;\quad r_0 = r(1 + u \delta r)\,,
\end{equation}
and we obtain the following edges and vertices in the diagrammatic expansion:
\begin{displaymath}
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[thick] (-1, 0) -- (1, 0);
    \node[left] at (-1, 0) {$n_1$};
    \node[right] at (1, 0) {$n_2$};
\end{tikzpicture}
    \quad
    \frac{V}{n_1^2R^{-2} + r}\delta_{n_1 + n_2}
\quad\quad
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[densely dashed, thick] (-1, -1) -- (0, 0) -- (-1, 1);
    \draw[thick] (+1, -1) -- (0, 0) -- (+1, 1);
    \node[above right] at (1, 1) {$n_1$};
    \node[below right] at (1, -1) {$n_2$};
\end{tikzpicture}
    -\frac{um^2}{4V}\delta_{n_1 + n_2}
\end{displaymath}
\begin{displaymath}
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[densely dashed, thick] (-1, -1) -- (0, 0);
    \draw[thick] (0, 0) -- (-1, 1);
    \draw[thick] (+1, -1) -- (0, 0) -- (+1, 1);
    \node[above left] at (-1, 1) {$n_1$};
    \node[above right] at (1, 1) {$n_2$};
    \node[below right] at (1, -1) {$n_3$};
\end{tikzpicture}
    -\frac{um}{6V^2}\delta_{n_1 + n_2 + n_3}
\quad\quad
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[thick] (-1, -1) -- (1, 1);
    \draw[thick] (-1, 1) -- ( 1, -1);
    \node[above left] at (-1, 1) {$n_1$};
    \node[above right] at (1, 1) {$n_2$};
    \node[below right] at (1, -1) {$n_3$};
    \node[below left] at (-1, -1) {$n_4$};
\end{tikzpicture}
    -\frac{u}{24V^3}\delta_{n_1 + n_2 + n_3 + n_4}
\end{displaymath}
plus additional vertices associated with the counterterms $\delta r$, $\delta
u$ which we do not list for brevity.

The logarithm of the probability distribution of $m$ is the sum of all
connected diagrams:
\begin{displaymath}
\begin{split}
    \log p(m) = &\,
    \mathrm{const} -V \left(\frac{1}{2} r m^2 + 
    \frac{u}{24} m^4\right)\\
&+\ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-1.75, -0.75) -- (-1, 0) -- (-1.75, 0.75);
    \draw[thick] (-1, 0) .. controls (-0.2, 0.8) and (0.5, 0.8) .. (0.5, 0) .. 
    controls (0.5, -0.8) and (-0.2, -0.8) .. (-1, 0);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-1.75, 0) -- (-0.75, 0);
    \draw[thick] (-0.75, 0) -- (0.75, 0);
    \draw[thick, densely dashed] (1.75, 0) -- (0.75, 0);
    \draw[thick] (0, 0) circle (0.75cm);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-2.75, -0.75) -- (-2, 0) -- (-2.75, 0.75);
    \draw[thick] (-2, 0) .. controls (-1.2, 0.8) and (-0.8, 0.8) .. (0, 0) .. 
    controls (-0.8, -0.8) and (-1.2, -0.8) .. (-2, 0);
    \draw[thick] (1.6, 0) .. controls (1.6, 0.8) and (0.8, 0.8) .. (0, 0) .. 
    controls (0.8, -0.8) and (1.6, -0.8) .. (1.6, 0);
\end{tikzpicture}
 \ + \ \cdots \\
&+\ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-1.75, -0.75) -- (-1, 0) -- (-1.75, 0.75);
    \draw[thick, densely dashed] (1.75, -0.75) -- (1, 0) -- (1.75, 0.75);
    \draw[thick] (-1, 0) .. controls (-0.2, 0.8) and (0.2, 0.8) .. (1, 0) .. 
    controls (0.2, -0.8) and (-0.2, -0.8) .. (-1, 0);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-1.5, 0.75) -- (-0.75, 0) -- (-1.5, -0.75) ;
    \draw[thick] (0, 0) circle (0.75cm);
    \draw[thick, densely dashed] (0.53, 0.53) -- (1.25, 1.25);
    \draw[thick, densely dashed] (0.53, -0.53) -- (1.25, -1.25);
    \draw[thick] (0.53, -0.53) .. controls (0.2, -0.2) 
    and (0.2, 0.2) .. (0.53, 0.53);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-2.75, -0.75) -- (-2, 0) -- (-2.75, 0.75);
    \draw[thick, densely dashed] (2.75, -0.75) -- (2, 0) -- (2.75, 0.75);
    \draw[thick] (-2, 0) .. controls (-1.2, 0.8) and (-0.8, 0.8) .. (0, 0) .. 
    controls (-0.8, -0.8) and (-1.2, -0.8) .. (-2, 0);
    \draw[thick] (2, 0) .. controls (1.2, 0.8) and (0.8, 0.8) .. (0, 0) .. 
    controls (0.8, -0.8) and (1.2, -0.8) .. (2, 0);
\end{tikzpicture}
 \ + \ \cdots \\
& + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, densely dashed] (-1.5, 0.75) -- (-0.75, 0) -- (-1.5, -0.75) ;
    \draw[thick] (0, 0) circle (0.75cm);
    \draw[thick, densely dashed] (1.40, -0.92) -- (0.375, -0.65) -- (0.1, -1.67);
    \draw[thick, densely dashed] (1.40, 0.92) -- (0.375, 0.65) -- (0.1, 1.67);
\end{tikzpicture}
 \ + \ \cdots \,,
\end{split}
\end{displaymath}
where again we left out all diagrams involving the counterterms for brevity.

Retaining only one-loop diagrams, and employing a heath kernel regulator we have:
\begin{equation}
\begin{split}
    -\log p(m) = \mathrm{const} + V \Bigg( 
    &\frac{1}{2} m^2\left( r + u \left(\delta r + \frac{1}{2} I_1\right) + O(u^2)\right) + \\
    &\frac{u}{24} m^4 \left(1 + u \left(\delta u - \frac{3}{2} I_2\right) + O(u^2)\right) + \\
    &\frac{u^3}{48}m^6\left(I_3 + O(u)\right) + O(u^4 m^8) \Bigg)\,,
\end{split}
\end{equation}
where
\begin{equation}
    I_k = \frac{1}{V} \sum_{n\in\mathbb{Z}^4\setminus\{0\}} 
    \frac{e^{-s(n^2 R^{-2} + r)}}{\left(n^2 R^{-2} + r\right)^k} = 
    \frac{1}{(k - 1)!}\left(-\frac{\partial}{\partial r} - s\right)^{k-1} I_1
\end{equation}
